---
#output: html_document
output:
  github_document:
    html_preview: FALSE
  html_preview: TRUE
params:
  prepared_by:
    label: "Prepared by:"
    value: "Richard Martin-Nielsen"
    input: text
title: "Sub-national COVID graphs for Lithuania"
---


```{r setup, include=FALSE}

library(covidregionaldata)
library(ggplot2)
library(ggridges)
library(roll)
library(scales)
library(forcats)
library(dplyr)
library(tidyr)

country <- "Lithuania"
# Set-up output ----
# Figure path on disk = base.dir + fig.path
# Figure URL online = base.url + fig.path
knitr::opts_knit$set(base.dir = stringr::str_c(here::here(), "/docs/"), base.url = "/covidregionaldatagraphs/") # project root folder
knitr::opts_chunk$set(fig.path = stringr::str_c(paste0(paste("images", country, sep="/")), "-"))

dataset_details <- get_available_datasets("regional") %>% filter(grepl(country, origin))

lt_county_data <- get_regional_data(
  country = "Lithuania",
  totals = FALSE,
  level = 1,
  localise = TRUE,
  all_osp_fields = TRUE
)

lt_municipality_data <- get_regional_data(
  country = "Lithuania",
  totals = FALSE,
  level = 2,
  localise = TRUE,
  all_osp_fields = TRUE,
  national_data = TRUE
)

lt_national_data <- lt_municipality_data %>%
  filter(municipality == "Lietuva") %>%
  select(-municipality, -iso_3166_2_municipality, -county, -iso_3166_2)

lt_municipality_data <- lt_municipality_data %>%
  filter(municipality != "Lietuva")

lt_last_date <- format(max(lt_municipality_data$date), "%b %d, %Y")

caption_text <- paste0(
  "Data: ",
  dataset_details$source_text,
  " (sourced through covidregionaldata), ",
  lt_last_date, 
  "\nPrepared by: ",
  params$prepared_by
)
# Make summary table ----

region_summaries <-
  lt_municipality_data %>%
  group_by(municipality) %>%
  summarise(
    min_i = min(cases_new), max_i = max(cases_new),
    median_i = median(cases_new), mean_i = mean(cases_new)
  ) %>%
  rename(region = municipality)

narrowed_regions <- pull(region_summaries %>% slice_max(max_i, n = 10) %>% select(region))

narrowed_regional_incidence <- lt_municipality_data %>%
  # filter(date < as_date("2021-01-04")) %>%
  filter(municipality %in% narrowed_regions)

# Set graphing defaults ----

theme_set(
  theme_minimal() +
    theme(plot.caption = element_text(size = 8))
)

```

# COVID-19 Data Charts for Lithuania

This page presents a set of static charts representing
data on the COVID-19 pandemic in Lithuania. These charts show only past data.

This code uses the
[`covidregionaldata`](http://epiforecasts.io/covidregionaldata) package
to download data for various countries at a subregional level. The workflow
is currently set to install the github version of the package.

This page has some additional graphs specific to Lithuania, in addition
to those available in the other
[page of graphs showing COVID-19 data for Lithuania](COVID-19 regional graphs - Lithuania.html)

## Overall comparisons of municipalities

Two charts imitating those prepared by the OSP in [their analyses](https://osp.stat.gov.lt/documents/10180/8420714/1_COVID-19_situacijos_apzvalga_210215.pdf).

```{r waterfall-charts, echo=FALSE}
# Waterfall chart case counts ----
lt_counts <- lt_municipality_data %>%
  select(date, cases_new, municipality) %>%
  mutate(weekly_mean_cases = roll_mean(cases_new, 7)) %>%
  filter(date > "2020-10-01") %>%
  group_by(date,
    group = fct_rev(cut(weekly_mean_cases,
      breaks = c(-1, 0, 5, 10, 15, 20, Inf),
      labels = c("0", "0-5", "5-10", "10-15", "15-20", "20+"),
      include.lowest = TRUE
    ))
  ) %>%
  summarise(count = n(), .groups = "drop_last")

lt_counts %>%
  ggplot() +
  geom_col(mapping = aes(x = date, y = count, fill = group), width = 1) +
  labs(
    x = "Date",
    y = "Number of municipalities",
    fill = "Case count",
    title = "Municipality case counts in Lithuania",
    subtitle = "7 day average counts of new cases",
    caption = caption_text
  ) +
  scale_fill_brewer(palette = "Blues", direction = 1)

#ggsave("extra/Lithuania-waterfall-case-counts.png", width = 6, height = 4, units = "in")

# Waterfall chart municipality test positivity ----
lt_positivity <- lt_municipality_data %>%
  filter(date > "2020-10-01") %>%
  select(date, dgn_prc_day, municipality) %>%
  mutate(weekly_mean_positivity = roll_mean(dgn_prc_day, 7)) %>%
  group_by(date,
    group = fct_rev(cut(dgn_prc_day,
      breaks = c(-1, 0, 5, 10, 15, 20, Inf),
      labels = c("0%", "0-5%", "5-10%", "10-15%", "15-20%", "20%+"),
      include.lowest = TRUE
    ))
  ) %>%
  summarise(count = n(), .groups = "drop_last")

lt_positivity %>%
  ggplot() +
  geom_col(mapping = aes(x = date, y = count, fill = group), width = 1) +
  labs(
    x = "Date",
    y = "Number of municipalities",
    fill = "Test positivity",
    title = "Municipality test positivity in Lithuania",
    subtitle = "7 day average test positivity",
    caption = caption_text
  ) +
  scale_fill_brewer(palette = "Oranges", direction = 1)

#ggsave("extra/Lithuania-waterfall-positivity.png", width = 6, height = 4, units = "in")

```

## Incidence charts

Ridgeline charts showing the incidence in regions of the country. These
emphasise that the first wave was barely a ripple compared with the
second wave, and that the incidence in the larger cities (particularly
Vilnius) has been quite separate from other municipalities.
                                                          
```{r incidence-ridgeline-charts, echo=FALSE}
# Plot ridgeline incidence for top 10 municipalities ----

intercity_gap <- 400

ridgeline_labels <- narrowed_regional_incidence %>%
  mutate(y = -as.numeric(factor(municipality)) * intercity_gap, region = municipality) %>%
  select(region, y) %>%
  unique()

narrowed_regional_incidence %>%
  ggplot(aes(
    x = date, y = -as.numeric(factor(municipality)) * intercity_gap,
    height = cases_new, group = municipality
  )) + # y=as.numeric(municipality)*250
  geom_ridgeline(alpha = 0.5, aes(fill = municipality), size = 0.25) +
  scale_y_continuous(
    breaks = ridgeline_labels$y,
    labels = ridgeline_labels$region,
    sec.axis =
      sec_axis(~ . + 10,
        name = "Daily incidence (confirmed cases)",
        labels = rep(c(intercity_gap / 2, "0"), 10),
        breaks = seq(
          from = -intercity_gap / 2,
          to = -intercity_gap * 10,
          by = -intercity_gap / 2
        )
      ),
    name = "Region"
  ) +
  scale_x_date(date_breaks = "3 months", date_minor_breaks = "1 month", date_labels = "%b") +
  theme_ridges() +
  theme(legend.position = "none") +
  labs(
    x = "Date", y = "Region / Incidence",
    title = "Regional COVID-19 incidence in Lithuania",
    subtitle = "Top ten municipalities by maximum daily incidence",
    caption = caption_text
  ) +
  theme(
    axis.text.y = element_text(size = 8),
    axis.title.y.left = element_blank()
  )

#ggsave("extra/Lithuania-ridgeline-top-municipalities.png", width = 6, height = 4, units = "in")

# Plot ridgeline incidence for all 10 counties ----

intercity_gap <- 800

ridgeline_labels <- lt_county_data %>%
  filter(county != "Unknown") %>%
  mutate(y = -as.numeric(factor(county)) * intercity_gap, region = county) %>%
  select(region, y) %>%
  unique()

lt_county_data %>%
  filter(county != "Unknown") %>%
  ggplot(aes(
    x = date, y = -as.numeric(factor(county)) * intercity_gap,
    height = cases_new, group = county
  )) + # y=as.numeric(municipality)*250
  geom_ridgeline(alpha = 0.5, aes(fill = county), size = 0.25) +
  scale_y_continuous(
    breaks = ridgeline_labels$y,
    labels = ridgeline_labels$region,
    sec.axis =
      sec_axis(~ . + 10,
        name = "Daily incidence (confirmed cases)",
        labels = rep(c(intercity_gap / 2, "0"), 10),
        breaks = seq(
          from = -intercity_gap / 2,
          to = -intercity_gap * 10,
          by = -intercity_gap / 2
        )
      ),
    name = "Region"
  ) +
  scale_x_date(date_breaks = "3 months", date_minor_breaks = "1 month", date_labels = "%b") +
  theme_ridges() +
  theme(legend.position = "none") +
  labs(
    x = "Date", y = "Region / Incidence",
    title = "Regional COVID-19 incidence in Lithuania",
    subtitle = "All counties",
    caption = caption_text
  ) +
  theme(
    axis.text.y = element_text(size = 8),
    axis.title.y.left = element_blank()
  )
#ggsave("extra/Lithuania-ridgeline-all-counties.png", width = 6, height = 4, units = "in")

```

The Lithuanian word for "municipality" is *savivaldybė*. "m. sav." is short
for "city municipality" and "r. sav." means "regional municipality"; Vilnius
and Kaunas both have suburban municipalities which are distinguished in these
charts only by the "m." and "r."


## Acceleration of case incidence and test positivity nationwide

Acceleration is another measure used by OSP. It is intended to give a
more responsive indicator of the development of the pandemic, particularly
in comparison with 7- and 14-day averages. The time-averaged values
are used to smooth out regular weekly variation in the data but are
necessarily lagging indicators, both when incidence is rising and when
it is declining.

```{r acceleration-national, echo=FALSE}

# Acceleration calculations - national ----

lt_national_data %>%
  # put rolling 7 day average in here
  mutate(
    weekly_mean_cases = roll_mean(cases_new, 7),
    weekly_mean_positivity = roll_mean(dgn_prc_day, 7)
  ) %>%
  mutate(
    cases_accel = ((weekly_mean_cases - lag(weekly_mean_cases)) / abs(lag(weekly_mean_cases))),
    test_accel = ((weekly_mean_positivity - lag(weekly_mean_positivity)) / abs(lag(weekly_mean_positivity)))
  ) %>%
  filter(date > "2020-09-01") %>%
  select(date, cases_accel, test_accel) %>%
  pivot_longer(
    cols = ends_with("_accel"),
    values_to = "accel",
    names_to = "type", names_pattern = "(.*)_accel"
  ) %>%
  mutate(type = if_else(type == "test", "test positivity", type)) %>%
  ggplot(aes(x = date, y = accel, colour = type)) +
  geom_line() +
  scale_x_date(date_breaks = "2 months", date_minor_breaks = "1 month", date_labels = "%B") +
  # scale_y_continuous(trans = modulus_trans(-0.5), labels=label_percent()) +
  scale_y_continuous(labels = label_percent()) +
  geom_hline(yintercept = 0, size = 0.2) +
  labs(
    x = "Date", y = "Acceleration",
    title = "Acceleration of the COVID-19 pandemic in Lithuania",
    subtitle = "% change in 7-day average of incidence or test positivity",
    caption = caption_text
  )


```

## Attributions of deaths

The OSP provides counts for three different criteria attributing deaths to
COVID, which roughly correspond to "of", "with" and "after". This chart
provides a comparison of how the three number of deaths attributed to COVID
according to each criterion compare.


```{r death-definitions, echo=FALSE}
# Compare death definition counts

lt_national_data %>%
  ggplot(aes(x = date)) +
  geom_col(aes(y = daily_deaths_def3), color = "grey70", width = 1) +
  geom_col(aes(y = daily_deaths_def2), color = "grey50", width = 1) +
  geom_col(aes(y = daily_deaths_def1), color = "black", width = 1) +
  labs(
    x = "Date", y = "Deaths",
    title = "Different counts of deaths due to COVID in Lithuania",
    subtitle = "Daily deaths 'by', 'with' or 'after' COVID",
    caption = caption_text
  )
```
